# Results

```{r}
library("tidyverse")
library(stringr)
library(lubridate)
library(plotly)

raw_data = read.csv("~/group24-EDAV-Project/data/allegations.csv")
raw_data$board_disposition_top = word(raw_data$board_disposition, 1) 
raw_data$year_month_received <- ymd(with(raw_data, sprintf("%d-%02d-%d", year_received, month_received, 28)))

raw_data <- raw_data%>%
  mutate(complainant_ethnicity=replace(complainant_ethnicity, complainant_ethnicity=="", "Unknown"))
```


## Mosaic

### Board Disposition vs Complaint Type

```{r fig.width=9, fig.height=8}
library("vcd")


vcd::mosaic(board_disposition_top ~ fado_type, raw_data, 
            direction = c("v", "h"), 
            highlighting_fill = c("grey90", "cornflowerblue", "coral2"),
            abbreviate_labs  = c(15, 15),
            labeling_args = list(set_varnames = c(fado_type="Type", 
                                                  board_disposition_top="Board Disposition")))
```

### Complaint Type vs Complainant Ethnicity


```{r fig.width=9, fig.height=8}

top_3_complainant_ethnicity <- raw_data %>% 
        group_by(complainant_ethnicity) %>% 
        summarize(count=n()) %>% 
        arrange(desc(count)) %>% 
        top_n(4)

top_3_complainant_ethnicity_filtered <- raw_data %>% 
                      filter(complainant_ethnicity %in% top_3_complainant_ethnicity$complainant_ethnicity)


vcd::mosaic(fado_type ~ complainant_ethnicity, top_3_complainant_ethnicity_filtered, 
            direction = c("v", "h"), 
            highlighting_fill = c("coral2","grey90", "cornflowerblue", "darkblue"),
            abbreviate_labs  = c(15, 15),
            labeling_args = list(set_varnames = c(complainant_ethnicity="Complainant Ethnicity", 
                                                  board_disposition_top="Board Disposition")))
```


### Board Disposition vs Complainant Ethnicity


```{r fig.width=9, fig.height=8}

top_3_complainant_ethnicity <- raw_data %>% 
        filter(year_received > 2000) %>% 
        group_by(complainant_ethnicity) %>% 
        summarize(count=n()) %>% 
        arrange(desc(count)) %>% 
        top_n(4)

top_3_complainant_ethnicity_filtered <- raw_data %>% 
                      filter(complainant_ethnicity %in% top_3_complainant_ethnicity$complainant_ethnicity)


vcd::mosaic(board_disposition_top ~ complainant_ethnicity, top_3_complainant_ethnicity_filtered, 
            direction = c("v", "h"), 
            highlighting_fill = c("grey90", "cornflowerblue", "coral2"),
            abbreviate_labs  = c(15, 15),
            labeling_args = list(set_varnames = c(complainant_ethnicity="Complainant Ethnicity", 
                                                  board_disposition_top="Board Disposition")))
```
### Board Disposition vs Complainant Gender


```{r fig.width=9, fig.height=8}

top_3_complainant_gender <- raw_data %>% 
        group_by(complainant_gender) %>% 
        summarize(count=n()) %>% 
        arrange(desc(count)) %>% 
        top_n(2)

top_3_complainant_gender_filtered <- raw_data %>% 
                      filter(complainant_gender %in% top_3_complainant_gender$complainant_gender)


vcd::mosaic(board_disposition_top ~ complainant_gender, top_3_complainant_gender_filtered, 
            direction = c("v", "h"), 
            highlighting_fill = c("grey90", "cornflowerblue", "coral2"),
            abbreviate_labs  = c(15, 15),
            labeling_args = list(set_varnames = c(complainant_gender="Complainant Gender", 
                                                  board_disposition_top="Board Disposition")))
```


```{r fig.width=9, fig.height=8}

top_3_mos_ethnicity <- raw_data %>% 
        group_by(mos_ethnicity) %>% 
        summarize(count=n()) %>% 
        arrange(desc(count)) %>% 
        top_n(4)

top_3_mos_ethnicity_filtered <- raw_data %>% 
                      filter(mos_ethnicity %in% top_3_mos_ethnicity$mos_ethnicity) 


vcd::mosaic(board_disposition_top ~ mos_ethnicity, top_3_mos_ethnicity_filtered, 
            direction = c("v", "h"), 
            highlighting_fill = c("grey90", "cornflowerblue", "coral2"),
            abbreviate_labs  = c(15, 15),
            labeling_args = list(set_varnames = c(mos_ethnicity="Officer's Ethinicity ", 
                                                  board_disposition_top="Board Disposition")))

top_3_mos_ethnicity_filtered <- raw_data %>% 
                      filter(mos_ethnicity %in% top_3_mos_ethnicity$mos_ethnicity) %>% 
                      group_by(mos_ethnicity, board_disposition_top) %>% 
                      summarize(count=n())

p <- ggplot(top_3_mos_ethnicity_filtered, aes(x = mos_ethnicity, y = count))+
  geom_col(aes(fill = board_disposition_top), width = 0.7)+ 
  xlab("Officer's Ethinicity") + 
  ylab("Case Count") + 
  guides(fill=guide_legend(title="Board Disposition"))
p
```


```{r}



count_by_individual <- raw_data %>% 
  filter(mos_ethnicity %in% top_3_mos_ethnicity$mos_ethnicity) %>% 
  group_by(unique_mos_id, first_name, last_name, rank_now, mos_ethnicity, mos_gender) %>% 
  summarize(count=n())


p <- ggplot(count_by_individual, aes(x = fct_reorder(mos_ethnicity, count, median), y = count))+ 
  geom_boxplot() +
  coord_flip() +
  ylab("No of Cases per Officer") +
  xlab("Officer's Ethinicity") +
  theme_minimal() 

p
```

```{r}
library(forcats)
p <- ggplot(count_by_individual, aes(x = fct_reorder(rank_now, count, median), y = count))+ 
  geom_boxplot() +
  coord_flip() +
  ylab("No of Cases per Officer") +
  xlab("Officer's Ethinicity") +
  theme_gray() 

p

```


```{r fig.width=7, fig.height=6}

count_by_individual <- raw_data %>% 
  filter(mos_ethnicity %in% top_3_mos_ethnicity$mos_ethnicity) %>% 
  group_by(unique_mos_id, first_name, last_name, rank_now, mos_ethnicity, mos_gender, board_disposition_top) %>% 
  summarize(count=n())


p <- ggplot(count_by_individual, aes(x = fct_reorder(mos_ethnicity, count, median), y = count))+ 
  geom_boxplot() +
  coord_flip() +
  ylab("No of Cases per Officer") +
  xlab("Officer's Ethinicity") +
  theme_grey(16) 

p + facet_grid(board_disposition_top ~ .)

```

```{r}
#Cleaveland Dot Plot Theme

theme_dotplot <- theme_bw(14) +
  theme(axis.text.y = element_text(size = rel(.75)),
        axis.ticks.y = element_blank(),
        axis.title.x = element_text(size = rel(.75)),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(size = 0.5),
        panel.grid.minor.x = element_blank())
```

```{r}
count_by_allegation <- raw_data %>% 
  group_by(fado_type, allegation) %>% 
  summarise(count = n(), .groups="drop") %>% 
  arrange(desc(count))
```

## Force

An officer used excessive or otherwise unnecessary force.

```{r fig.width=7, fig.height=6}
# create the plot
count_by_allegation %>% 
  filter(fado_type == "Force") %>% 
ggplot(aes(x = count, y = reorder(allegation, count))) +
  geom_point(color = "blue") +
  scale_x_continuous(limits = c(0, 5000)) +
  theme_dotplot +
  xlab("No of Allegations Made ") +
  ylab("Allegation\n") +
  ggtitle("Force")
```

A total of {}

## Abuse of Authority 

An officer used police powers to take unwarranted actions, such as an unlawful search.

```{r fig.width=7, fig.height=8}
# create the plot
count_by_allegation %>% 
  filter(fado_type == "Abuse of Authority") %>% 
ggplot(aes(x = count, y = reorder(allegation, count))) +
  geom_point(color = "blue") +
  scale_x_continuous(limits = c(0, 2500)) +
  theme_dotplot +
  xlab("No of Allegations Made") +
  ylab("Allegations\n") +
  ggtitle("Abuse of Authority")
```


## Discourtesy

An officer engaged in rude or profane behavior toward a civilian.
```{r fig.width=7, fig.height=3}
# create the plot
count_by_allegation %>% 
  filter(fado_type == "Discourtesy") %>% 
ggplot(aes(x = count, y = reorder(allegation, count))) +
  geom_point(color = "blue") +
  scale_x_continuous(limits = c(0, 4000)) +
  theme_dotplot +
  xlab("No of Allegations Made") +
  ylab("Allegations\n") +
  ggtitle("Discourtesy")
```

## Offensive Language

An officer used one or more slurs relating to race, ethnicity, gender, religion, sexual orientation or disability.

```{r fig.width=7, fig.height=4}
# create the plot
count_by_allegation %>% 
  filter(fado_type == "Offensive Language") %>% 
ggplot(aes(x = count, y = reorder(allegation, count))) +
  geom_point(color = "blue") +
  scale_x_continuous(limits = c(0, 400)) +
  theme_dotplot +
  xlab("No of Allegations Made") +
  ylab("Allegations\n") +
  ggtitle("Offensive Language")
```


## Time series analysis

```{r}

complaints_by_time <- raw_data %>% 
  filter(year_received < 2020) %>% 
  group_by(year_received, fado_type) %>% 
  summarize(count = n())


g <- ggplot(complaints_by_time, aes(year_received, count,  color = fado_type)) + 
  geom_line() +
  xlab("Year") +
  ylab("No of Cases Received")+ 
  guides(color=guide_legend(title="Complaint Type"))

ggplotly(g)
```



```{r}

complaints_by_time <- raw_data %>% 
  filter(year_received < 2020) %>% 
  group_by(year_received, mos_ethnicity) %>% 
  summarize(count = n())


g <- ggplot(complaints_by_time, aes(year_received, count,  color = mos_ethnicity)) + 
  geom_line() +
  xlab("Year") +
  ylab("No of Cases Received")+ 
  guides(color=guide_legend(title="Officer Ethnicity"))

ggplotly(g)
```



```{r}

complaints_by_time <- raw_data %>% 
  filter(year_received < 2020) %>% 
  group_by(year_received, complainant_ethnicity) %>% 
  summarize(count = n())


g <- ggplot(complaints_by_time, aes(year_received, count,  color = complainant_ethnicity)) + 
  geom_line() +
  xlab("Year") +
  ylab("No of Cases Received")+ 
  guides(color=guide_legend(title="Complainant Ethnicity"))

ggplotly(g)
```


```{r}

complaints_by_time <- raw_data %>% 
  filter(year_received < 2020) %>% 
  filter(complainant_gender %in% top_3_complainant_gender$complainant_gender) %>% 
  group_by(year_received, complainant_gender) %>% 
  summarize(count = n())


g <- ggplot(complaints_by_time, aes(year_received, count,  color = complainant_gender)) + 
  geom_line() +
  xlab("Year") +
  ylab("No of Cases Received")+ 
  guides(color=guide_legend(title="Complaint Gender"))

ggplotly(g)
```

```{r}

complaints_by_time <- raw_data %>% 
  filter(year_received < 2020) %>% 
  group_by(year_received, board_disposition_top) %>% 
  summarize(count = n())


g <- ggplot(complaints_by_time, aes(year_received, count,  color = board_disposition_top)) + 
  geom_line() +
  xlab("Year") +
  ylab("No of Cases Received")+ 
  guides(color=guide_legend(title="Board Disposition"))

ggplotly(g)
```


```{r}

complaints_by_time <- raw_data %>% 
  filter(year_received < 2020, year_received > 1995) %>% 
  group_by(year_received, complainant_ethnicity, board_disposition_top) %>% 
  summarize(count = n())


g <- ggplot(complaints_by_time, aes(year_received, count,  color = complainant_ethnicity)) + 
  geom_line() +
  xlab("Year") +
  ylab("No of Cases Received")+ 
  guides(color=guide_legend(title="Complainant Ethnicity"))

ggplotly(g + facet_grid(board_disposition_top ~ .))
```

```{r}
complaints_by_time <- raw_data %>% 
  filter(year_closed > 1995) %>%
  group_by(year_closed, complainant_ethnicity, board_disposition_top) %>% 
  summarize(count = n())


g <- ggplot(complaints_by_time, aes(year_closed, count,  color = complainant_ethnicity)) + 
  geom_line() +
  xlab("Year") +
  ylab("No of Cases Received")+ 
  guides(color=guide_legend(title="Complainant Ethnicity"))

ggplotly(g + facet_grid(board_disposition_top ~ .))
```